PeerMind

### 2016-08-12 paper.tex
Fixing delegations.

\begin{displaymath}
\begin{array}{rcl}
V_{M_0} & = & 0 \\
V_{M_1} & = & r_{1,0} V_{M_0} + r_{1,3} V_{M_3} = 0.75 V_{M_0} + 0.25 V_{M_3} \\
V_{M_2} & = & r_{2,3} V_{M_3} = 1.00 V_{M_3} \\
V_{M_3} & = & r_{3,1} V_{M_1} + r_{3,2} V_{M_2} + r_{3,4} V_{M_4} = \\
& = & 0.40 V_{M_1} + 0.40 V_{M_2} + 0.20 V_{M_4} \\
\end{array}
\end{displaymath}

@@

We solve the matrix equation:

\begin{displaymath}
\mathrm{D} \mathbf{x_v} = \mathbf{v}
\end{displaymath}

@@

\begin{displaymath}
\mathbf{v} = \left[ \begin{array}{c}
0.7 \\
0 \\
-0.4 \\
0 \\
0 \\
\end{array} \right],\quad \mathbf{x_v} = \left[ \begin{array}{c}
0.7 \\
0.54 \\
-0.4 \\
0.056 \\
0.0 \\
\end{array} \right]
\end{displaymath}

There are few details we have to look into.
Despite delegation some votes or parts of votes can still happen to be discarded and should count
towards non-voting population.
Members might not declare their delegates.
Other members can delegate to them as well, potentially propagating lack of a vote further.
In our example, member $M_4$ has not voted so their vote is discarded.
As a consequence $M_3$'s vote cannot be computed fully, which influences $M_1$'s vote as well.

In such case we normalize ratios of other delegates, those who voted or for which we can
compute their votes recursively.
So $r_{3,1}$ and $r_{3,2}$ are normalized from $0.40$ to $0.50$.

Another issue is a question of abstentions.
In PeerMind members can abstain from voting which is different from simply not participating.
Abstentions are counted towards quorum, but they do not contribute to the result.
To address this, we define two vectors:

@@

We solve $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$.
To normalize votes we compute $x = x_v / x_w$.
A consequence is that those elements of $x_w$ equal to $0$ make corresponding elements of $x$ undefined.
Those undefined elements of $x$ we count towards non-voting population.
We just have to make sure that they are not in fact abstentions.
We check corresponding elements of $x_a$ and if they are non-zero we count them as abstentions,
otherwise as non-votes.
In our example:

\begin{displaymath}
\mathbf{x} = \left[ \begin{array}{c}
0.7 \\
0.57 \\
-0.4 \\
0.07 \\
\mathrm{undefined} \\
\end{array} \right]
\end{displaymath}

\todo[inline]{Are we discarding abstentions too much here? If we have 5 people and each of them is computed 0.2 of abstention
through delegation, while the rest 0.8 is a vote, then we normalize 0.8 to 1.0 for all of them and simply discard 0.2
of abstention. But otherwise those 0.2 of abstentions would count together towards 1.0. But probably this is not a
problem because the quorum is still the same: those 5 people have participated through delegation in voting. Only
if somebody has not participated through delegation in voting we care if they might participated in abstentions
through voting. To maximize delegated participation.}

The result (average of scores) without delegation is $\overline{v} = 0.15$ and
with delegation is $\overline{x} = 0.24$.
Delegation pushes decision slightly in favor.

\todo[inline]{$x = x_v / x_w$ does not really produce the same result as computing with 0.50 as ratio to begin
with. Why?}

>>>>

\begin{displaymath}
\begin{array}{rcl}
V_{M_0} & = & V_{M_0} \\
V_{M_1} & = & r_{1,0} V_{M_0} + r_{1,3} V_{M_3} = 0.75 V_{M_0} + 0.25 V_{M_3} \\
V_{M_2} & = & r_{2,3} V_{M_3} = 1.00 V_{M_3} \\
V_{M_3} & = & r_{3,1} V_{M_1} + r_{3,2} V_{M_2} + r_{3,4} V_{M_4} = \\
& = & 0.40 V_{M_1} + 0.40 V_{M_2} + 0.20 V_{M_4} \\
V_{M_4} & = & V_{M_4} \\
\end{array}
\end{displaymath}

@@

We solve the matrix equation for $\mathbf{x_v}$:

\begin{displaymath}
\mathrm{D} \mathbf{x_v} = \mathbf{v},\quad \mathbf{v} = \left[v_{i}\right] = \begin{dcases*}
 V_{M_i} & if $M_i$ voted \\
 0 & otherwise \\
\end{dcases*}
\end{displaymath}

@@

\begin{displaymath}
\mathbf{v} = \left[ \begin{array}{c}
0.70 \\
0 \\
-0.4 \\
0 \\
0 \\
\end{array} \right],\quad \mathbf{x_v} = \left[ \begin{array}{c}
0.70 \\
0.54 \\
-0.40 \\
0.056 \\
0 \\
\end{array} \right]
\end{displaymath}

There are few details we have to look into.
Despite delegation some votes or parts of votes can still happen to be discarded.
Members who did not vote might have not declared their delegates.
Other members could delegate to them as well, potentially propagating lack of a vote further.
In our example, member $M_4$ has not voted so their vote is discarded.
As a consequence $M_3$'s vote cannot be computed fully, which influences $M_1$'s vote as well.

In such case we want to normalize ratios of other delegates, those who voted or for which we can
compute their votes recursively.
So $r_{3,1}$ and $r_{3,2}$ are normalized from $0.40$ to $0.50$.

Another issue is a question of abstentions.
In PeerMind members can abstain from voting which is different from simply not participating.
Abstentions are counted towards quorum, but they do not contribute to the result.

Moreover, from $\mathbf{x_v}$ alone we cannot see which elements represent vote $0$ and which
represent that it was not possible to compute a vote for a member.

To address all this, we define two vectors:

@@

We solve $\mathrm{D} \mathbf{x_w} = \mathbf{w}$ and $\mathrm{D} \mathbf{x_a} = \mathbf{a}$ for
$\mathbf{x_w}$ and $\mathbf{x_a}$, respectively.
Non-zero elements of $\mathbf{x_w}$ correspond to members who voted or for which we can compute
a vote.
We define a normalized matrix:

\begin{displaymath}
\mathrm{D'} = \left[d'_{ij}\right] = \begin{dcases*}
 1 & if $i = j$ \\
 0 & if $x_{w_j} = 0$ ($V_{M_j}$ cannot be computed) \\
 -\frac{r_{i,j}}{R_i} & \parbox[t]{.55\columnwidth}{if $M_i$ did not vote, but delegates to $M_j$ with normalized ratio $\frac{r_{i,j}}{R_i}$, $R_i = \sum_{x_{w_k} \ne 0} r_{i,k}$} \\
 0 & otherwise \\
\end{dcases*}
\end{displaymath}

To compute votes with normalized ratios for only those delegates who voted or for which we can compute votes recursively,
we solve $\mathrm{D'} \mathbf{x} = \mathbf{v}$ for $\mathbf{x}$:

\begin{displaymath}
\mathbf{x} = \left[ \begin{array}{c}
0.70 \\
0.54 \\
-0.40 \\
0.07 \\
0 \\
\end{array} \right]
\end{displaymath}

If an element of $\mathbf{x_w}$ is equal to $0$, then we could not compute a vote for that member and we count
it towards non-voting population.
We just have to make sure that they are not in fact abstentions.
We check corresponding elements of $\mathbf{x_a}$ and if they are non-zero we count them as abstentions,
otherwise as non-votes.

The result (average of scores) without delegation is $\overline{v} = 0.15$ and
with delegation is $\overline{x} = 0.23$.
Delegation pushes decision slightly in favor.
